---
title: "Deprivation and multidimensional poverty"
subtitle: "Annual 2025 executive profile of income poverty, basic-needs deprivation and multidimensional poverty."
description: "Annual 2025 executive profile of income poverty, NBI and multidimensional poverty indicators for Ecuador, based on curated aggregate site outputs."
format:
  html:
    page-layout: full
    toc: false
    code-tools: false
    title-block-style: none
    body-class: epm-internal
---

```{=html}
<style>
.deprivation-hero {
  grid-template-columns: minmax(0, 1fr) minmax(340px, 400px);
  gap: clamp(2rem, 5vw, 4.5rem);
  overflow-x: clip;
}

.deprivation-hero .internal-hero-copy {
  min-width: 0;
  max-width: 760px;
}

.deprivation-hero h1 {
  font-size: clamp(3rem, 5.8vw, 5.2rem);
  line-height: 1;
  max-width: 720px;
  text-wrap: balance;
  overflow-wrap: normal;
}

.deprivation-hero .story-lede {
  font-size: clamp(1.18rem, 1.7vw, 1.5rem);
  max-width: 680px;
}

.deprivation-hero .internal-hero-copy > p {
  max-width: 690px;
}

.deprivation-scope-card {
  align-self: center;
  min-width: 0;
  padding: clamp(1.15rem, 2.2vw, 1.55rem);
}

.deprivation-scope-card h2 {
  font-size: 1.25rem;
  margin-bottom: 0.9rem;
}

.deprivation-scope-list {
  display: grid;
  gap: 0.72rem;
  margin: 0;
}

.deprivation-scope-list div {
  display: grid;
  gap: 0.2rem;
  grid-template-columns: minmax(5rem, 0.45fr) minmax(0, 1fr);
  align-items: baseline;
  border-top: 1px solid rgba(18, 48, 71, 0.1);
  padding-top: 0.72rem;
}

.deprivation-scope-list div:first-child {
  border-top: 0;
  padding-top: 0;
}

.deprivation-scope-list dt {
  color: var(--epm-muted);
  font-size: 0.76rem;
  font-weight: 760;
  letter-spacing: 0.06em;
  text-transform: uppercase;
}

.deprivation-scope-list dd {
  color: var(--epm-navy);
  font-size: 0.98rem;
  font-weight: 720;
  margin: 0;
}

@media (max-width: 900px) {
  .deprivation-hero {
    grid-template-columns: minmax(0, 1fr);
  }

  .deprivation-hero h1 {
    font-size: clamp(2.6rem, 14vw, 4.2rem);
  }

  .deprivation-scope-card {
    align-self: stretch;
    max-width: 100%;
  }
}

@media (max-width: 560px) {
  .deprivation-scope-list div {
    grid-template-columns: minmax(0, 1fr);
  }
}
</style>
```

```{r}
#| include: false
site_output_path <- file.path("data", "derived", "site", "site_deprivation_multidimensional_kpis.rds")
if (!file.exists(site_output_path)) {
  site_output_path <- file.path("..", "data", "derived", "site", "site_deprivation_multidimensional_kpis.rds")
}

kpis <- readRDS(site_output_path)
indicator_order <- c(
  "poverty_rate",
  "extreme_poverty_rate",
  "nbi_rate",
  "extreme_nbi_rate",
  "tpm_rate",
  "tpem_rate"
)
domain_order <- c("national", "urban", "rural")

pick_kpi <- function(target_indicator, target_domain) {
  rows <- subset(kpis, indicator_id == target_indicator & domain == target_domain)
  if (nrow(rows) == 0L) {
    return(NULL)
  }
  head(rows, 1L)
}

kpi_value <- function(indicator_id, domain) {
  row <- pick_kpi(indicator_id, domain)
  if (is.null(row)) {
    return(NA_real_)
  }
  as.numeric(row$display_estimate)
}

fmt_pct <- function(value) {
  if (is.na(value)) {
    return("Not available")
  }
  sprintf("%.1f%%", value)
}

fmt_num <- function(value) {
  if (is.na(value)) {
    return("n.a.")
  }
  format(round(value), big.mark = ",", scientific = FALSE)
}

indicator_label <- function(indicator_id) {
  switch(
    indicator_id,
    poverty_rate = "Income poverty",
    extreme_poverty_rate = "Extreme income poverty",
    nbi_rate = "NBI / Unsatisfied Basic Needs poverty",
    extreme_nbi_rate = "Extreme NBI",
    tpm_rate = "Multidimensional poverty / TPM",
    tpem_rate = "Extreme multidimensional poverty / TPEM",
    indicator_id
  )
}

indicator_short <- function(indicator_id) {
  switch(
    indicator_id,
    poverty_rate = "Income",
    extreme_poverty_rate = "Extreme income",
    nbi_rate = "NBI",
    extreme_nbi_rate = "Extreme NBI",
    tpm_rate = "TPM",
    tpem_rate = "TPEM",
    indicator_id
  )
}

indicator_family_label <- function(indicator_id) {
  switch(
    indicator_id,
    poverty_rate = "Monetary poverty",
    extreme_poverty_rate = "Monetary poverty",
    nbi_rate = "Basic-needs deprivation",
    extreme_nbi_rate = "Basic-needs deprivation",
    tpm_rate = "Multidimensional poverty",
    tpem_rate = "Multidimensional poverty",
    "Analytical indicator"
  )
}

indicator_note <- function(indicator_id) {
  switch(
    indicator_id,
    poverty_rate = "Household per-capita income below the poverty line.",
    extreme_poverty_rate = "Household per-capita income below the extreme poverty line.",
    nbi_rate = "People in households with at least one unsatisfied basic need.",
    extreme_nbi_rate = "People in households with two or more unsatisfied basic needs.",
    tpm_rate = "People with overlapping weighted deprivations above the TPM threshold.",
    tpem_rate = "People with severe overlapping weighted deprivations above the TPEM threshold.",
    "Survey-weighted analytical estimate."
  )
}

status_label <- function(status) {
  if (is.na(status) || !nzchar(status)) {
    return("Method status not available")
  }
  switch(
    status,
    annual_published_table_close_match = "Benchmark close match",
    not_directly_benchmarked_against_public_annual_tabulations = "Not directly benchmarked",
    benchmark_comparison_outside_tolerance = "Benchmark requires review",
    status
  )
}

domain_label <- function(domain) {
  switch(
    domain,
    national = "National",
    urban = "Urban",
    rural = "Rural",
    domain
  )
}

kpi_card <- function(indicator_id) {
  row <- pick_kpi(indicator_id, "national")
  if (is.null(row)) {
    return("")
  }
  paste0(
    '<div class="proof-item">',
    "<strong>", indicator_label(indicator_id), "</strong>",
    '<div style="font-size: 2rem; line-height: 1.05; font-weight: 800; margin: 0.35rem 0;">',
    fmt_pct(row$display_estimate),
    "</div>",
    "<p>", indicator_note(indicator_id), "</p>",
    '<p><strong>', status_label(row$benchmark_status), "</strong></p>",
    "</div>"
  )
}

table_cell <- function(indicator_id, domain) {
  value <- kpi_value(indicator_id, domain)
  paste0("<td>", fmt_pct(value), "</td>")
}

domain_table_row <- function(domain) {
  cells <- vapply(
    indicator_order,
    table_cell,
    character(1),
    domain = domain
  )
  paste0("<tr><th>", domain_label(domain), "</th>", paste(cells, collapse = ""), "</tr>")
}

rural_gap <- function(indicator_id) {
  kpi_value(indicator_id, "rural") - kpi_value(indicator_id, "urban")
}

rural_higher <- sum(vapply(
  indicator_order,
  function(indicator_id) kpi_value(indicator_id, "rural") > kpi_value(indicator_id, "urban"),
  logical(1)
), na.rm = TRUE)

gap_data <- data.frame(
  indicator_id = indicator_order,
  gap = vapply(indicator_order, rural_gap, numeric(1)),
  stringsAsFactors = FALSE
)
top_gap_row <- subset(gap_data, gap == max(gap, na.rm = TRUE))
top_gap_indicator <- as.character(head(top_gap_row$indicator_id, 1L))

income_national <- kpi_value("poverty_rate", "national")
nbi_national <- kpi_value("nbi_rate", "national")
tpm_national <- kpi_value("tpm_rate", "national")
tpem_national <- kpi_value("tpem_rate", "national")
tpm_rural_ratio <- kpi_value("tpm_rate", "rural") / kpi_value("tpm_rate", "urban")
weighted_n_national <- fmt_num(pick_kpi("poverty_rate", "national")$weighted_n)
ipm_benchmark_status <- status_label(pick_kpi("tpm_rate", "national")$benchmark_status)
```

::: {.internal-hero .internal-hero-wide .deprivation-hero}
::: {.internal-hero-copy}
::: {.section-kicker}
Annual 2025 executive profile
:::

# Deprivation & multidimensional poverty

::: {.story-lede}
Annual 2025 executive profile of income poverty, basic-needs deprivation and multidimensional poverty.
:::

This profile reads Ecuador's 2025 annual ENEMDU evidence across monetary poverty, structural basic-needs deprivation, and multidimensional poverty. Monetary poverty, basic-needs deprivation and multidimensional poverty capture different dimensions of welfare, so they should be read together rather than collapsed into one headline.
:::

::: {.internal-hero-card .deprivation-scope-card}
## Scope

<dl class="deprivation-scope-list">
<div>
<dt>Period</dt>
<dd>2025 annual ENEMDU</dd>
</div>
<div>
<dt>Source</dt>
<dd>Public ENEMDU annual 2025 microdata</dd>
</div>
<div>
<dt>Coverage</dt>
<dd>National, urban and rural</dd>
</div>
<div>
<dt>Engine</dt>
<dd>enemduR</dd>
</div>
<div>
<dt>Output</dt>
<dd>Curated aggregate RDS</dd>
</div>
<div>
<dt>Method</dt>
<dd>Survey-weighted analytical estimates</dd>
</div>
</dl>
:::
:::

::: {.content-block .editorial-flow}
::: {.story-panel}
::: {.section-kicker}
Executive signal
:::

## Three poverty lenses, one reading problem

The national income-poverty rate is **`r fmt_pct(income_national)`**, while NBI reaches **`r fmt_pct(nbi_national)`** and TPM reaches **`r fmt_pct(tpm_national)`**. These are not competing versions of the same statistic. Income poverty identifies insufficient current resources, NBI captures accumulated structural deficits, and TPM captures overlapping weighted deprivations.

The curated output represents about **`r weighted_n_national`** people nationally for the income-poverty denominator and stores only aggregate fields used by this page.
:::

::: {.proof-strip}
```{r}
#| echo: false
#| results: asis
cat(paste(vapply(indicator_order, kpi_card, character(1)), collapse = "\n"))
```
:::

::: {.infographic-frame}
::: {.section-kicker}
Domain comparison
:::

## National, urban and rural profile

```{r}
#| echo: false
#| results: asis
header_cells <- vapply(indicator_order, indicator_short, character(1))
header_html <- paste0("<th>", header_cells, "</th>", collapse = "")
body_html <- paste(vapply(domain_order, domain_table_row, character(1)), collapse = "\n")
cat(paste0(
  '<table class="table table-sm">',
  "<thead><tr><th>Domain</th>", header_html, "</tr></thead>",
  "<tbody>", body_html, "</tbody>",
  "</table>"
))
```

Values are percentages. Urban and rural estimates are shown only for supported annual domains and are not interpolated.
:::

::: {.story-grid}
::: {.story-panel}
::: {.section-kicker}
Insight 1
:::

## Rural disadvantage is systematic

Rural areas show the highest burden in **`r rural_higher` of 6** indicators displayed. The largest urban-rural gap appears in **`r indicator_label(top_gap_indicator)`**, which indicates that territorial inequality is not confined to monetary poverty.
:::

::: {.story-panel}
::: {.section-kicker}
Insight 2
:::

## Monetary and structural poverty diverge

The national NBI rate is **`r fmt_pct(nbi_national)`**, above the national income-poverty rate of **`r fmt_pct(income_national)`**. Reading both together separates insufficient current income from accumulated deficits in housing, services, schooling and household dependency conditions.
:::

::: {.story-panel}
::: {.section-kicker}
Insight 3
:::

## TPM captures overlapping deprivation

TPM reaches **`r fmt_pct(tpm_national)`** nationally and rural TPM is about **`r sprintf("%.1f", tpm_rural_ratio)`x** the urban rate. TPEM reaches **`r fmt_pct(tpem_national)`** nationally, showing the severe tail of overlapping deprivations.
:::

::: {.story-panel}
::: {.section-kicker}
Insight 4
:::

## Benchmark status matters

Income poverty is benchmarked against public annual tabulations and shows close-match status. NBI is not directly benchmarked against public annual tabulations in this build. TPM/TPEM are calculated with enemduR and their public IPM benchmark comparison is marked **`r ipm_benchmark_status`**, so this page treats them as analytical estimates with an explicit reproducibility flag.
:::
:::

::: {.method-note}
## Method note

The indicators are based on public ENEMDU annual 2025 microdata and built with `enemduR`. Estimates are survey-weighted. Benchmark comparison is reported only where a public annual reference is available. The comparison documents analytical reproducibility and does not imply official institutional validation.
:::

::: {.story-panel}
::: {.section-kicker}
Technical footer
:::

## Output and safety contract

| Item | Value |
|---|---|
| Curated output | `data/derived/site/site_deprivation_multidimensional_kpis.rds` |
| Local annual output | `data/derived/annual/annual_deprivation_multidimensional_poverty.rds` |
| Pipeline | `scripts/03_build_annual_core.R` then `scripts/04_build_site_metadata.R` |
| Stored data | Aggregates only |
| Build date field | Not stored in the curated output |
| Safety validation | No private paths, no raw identifiers, no dynamic timestamps |
:::

::: {.proof-strip}
::: {.proof-item}
**Income poverty**

Annual monetary poverty and extreme poverty.
:::

::: {.proof-item}
**NBI**

Basic-needs deprivation and extreme NBI.
:::

::: {.proof-item}
**TPM / TPEM**

Multidimensional poverty and severe multidimensional poverty.
:::

::: {.proof-item}
**Publication layer**

Static Quarto page from a curated aggregate RDS.
:::
:::

::: {.page-nav}
<a href="02_income-poverty.qmd">&lt;- Income poverty</a>
<a href="../index.qmd">Home -&gt;</a>
:::
:::
