---
title: "Labor market and poverty"
subtitle: "Employment conditions and household vulnerability in annual ENEMDU 2025"
description: "A public labor-market and poverty reading for Ecuador in 2025, based on curated aggregate site outputs."
format:
  html:
    page-layout: full
    toc: false
    code-tools: false
    title-block-style: none
    body-class: epm-internal
---

```{=html}
<style>
.labor-hero {
  grid-template-columns: minmax(0, 1fr) minmax(320px, 390px);
  gap: clamp(2rem, 5vw, 4.5rem);
}

.labor-scope-card {
  align-self: center;
  min-width: 0;
}

.labor-scope-list {
  display: grid;
  gap: 0.72rem;
  margin: 0;
}

.labor-scope-list div {
  border-top: 1px solid rgba(18, 48, 71, 0.1);
  display: grid;
  gap: 0.2rem;
  grid-template-columns: minmax(5.4rem, 0.42fr) minmax(0, 1fr);
  padding-top: 0.72rem;
}

.labor-scope-list div:first-child {
  border-top: 0;
  padding-top: 0;
}

.labor-scope-list dt {
  color: var(--epm-muted);
  font-size: 0.76rem;
  font-weight: 760;
  letter-spacing: 0.06em;
  text-transform: uppercase;
}

.labor-scope-list dd {
  color: var(--epm-navy);
  font-size: 0.98rem;
  font-weight: 720;
  margin: 0;
}

.labor-kpi-strip {
  display: grid;
  gap: 1px;
  grid-template-columns: repeat(4, minmax(0, 1fr));
  margin: 1.4rem 0 2rem;
  overflow: hidden;
  border-radius: 8px;
  box-shadow: var(--epm-shadow);
}

.labor-kpi {
  background: var(--epm-navy);
  border-top: 4px solid var(--epm-gold);
  color: #ffffff;
  min-height: 170px;
  padding: 1.2rem;
}

.labor-kpi:nth-child(2) {
  border-top-color: var(--epm-blue);
}

.labor-kpi:nth-child(3) {
  border-top-color: var(--epm-red);
}

.labor-kpi:nth-child(4) {
  border-top-color: #ffffff;
}

.labor-kpi strong {
  display: block;
  font-size: 1.08rem;
  line-height: 1.25;
}

.labor-kpi .labor-value {
  display: block;
  font-size: clamp(2rem, 4vw, 3.1rem);
  font-weight: 820;
  letter-spacing: 0;
  line-height: 1;
  margin: 0.55rem 0;
}

.labor-kpi p {
  color: rgba(255, 255, 255, 0.78);
  font-size: 1rem;
  line-height: 1.55;
  margin: 0;
}

.labor-bars {
  display: grid;
  gap: 0.9rem;
}

.labor-bar-row {
  display: grid;
  grid-template-columns: minmax(10rem, 14rem) 1fr 5rem;
  gap: 0.75rem;
  align-items: center;
}

.labor-bar-row span {
  color: var(--epm-ink-soft);
  font-weight: 650;
}

.labor-meter {
  background: #e9ecef;
  border-radius: 999px;
  height: 0.78rem;
  overflow: hidden;
}

.labor-meter span {
  background: var(--epm-blue);
  display: block;
  height: 100%;
}

.labor-bars .labor-bar-row:nth-child(2) .labor-meter span {
  background: var(--epm-gold);
}

.labor-bars .labor-bar-row:nth-child(3) .labor-meter span {
  background: var(--epm-red);
}

.labor-table-wrap {
  overflow-x: auto;
}

.labor-table-wrap table {
  margin-bottom: 0.35rem;
  min-width: 720px;
}

.labor-table-wrap th {
  color: var(--epm-navy);
  font-size: 0.86rem;
  letter-spacing: 0.03em;
  text-transform: uppercase;
}

.labor-table-wrap td,
.labor-table-wrap th {
  vertical-align: middle;
}

@media (max-width: 1100px) {
  .labor-hero,
  .labor-kpi-strip {
    grid-template-columns: 1fr;
  }
}

@media (max-width: 720px) {
  .labor-scope-list div,
  .labor-bar-row {
    grid-template-columns: minmax(0, 1fr);
  }

  .labor-bar-row strong {
    text-align: left;
  }
}
</style>
```

```{r}
#| include: false
site_output_path <- file.path("data", "derived", "site", "site_labor_poverty_profiles.rds")
if (!file.exists(site_output_path)) {
  site_output_path <- file.path("..", "data", "derived", "site", "site_labor_poverty_profiles.rds")
}

if (!file.exists(site_output_path)) {
  stop("Labor-market poverty data are not available.", call. = FALSE)
}

labor <- readRDS(site_output_path)

required_columns <- c(
  "period",
  "domain",
  "domain_value",
  "indicator_id",
  "labor_value",
  "labor_label",
  "display_estimate",
  "display_unit",
  "estimated_count"
)

missing_columns <- setdiff(required_columns, names(labor))
if (length(missing_columns) > 0L) {
  stop("Labor-market poverty data are missing expected public fields.", call. = FALSE)
}

pick_row <- function(target_indicator, target_domain = "national", target_value = NULL) {
  rows <- labor[
    labor$indicator_id == target_indicator &
      labor$domain_value == target_domain,
    ,
    drop = FALSE
  ]

  if (!is.null(target_value)) {
    rows <- rows[rows$labor_value == target_value, , drop = FALSE]
  }

  if (nrow(rows) == 0L) {
    return(NULL)
  }

  rows[1L, , drop = FALSE]
}

fmt_pct <- function(value, digits = 1L) {
  if (is.na(value)) {
    return("n.a.")
  }
  sprintf(paste0("%.", digits, "f%%"), as.numeric(value))
}

fmt_count <- function(value) {
  if (is.na(value)) {
    return("n.a.")
  }
  format(round(as.numeric(value)), big.mark = ",", scientific = FALSE)
}

html_escape <- function(value) {
  value <- as.character(value)
  value <- gsub("&", "&amp;", value, fixed = TRUE)
  value <- gsub("<", "&lt;", value, fixed = TRUE)
  value <- gsub(">", "&gt;", value, fixed = TRUE)
  value <- gsub('"', "&quot;", value, fixed = TRUE)
  value
}

public_label <- function(value) {
  switch(
    as.character(value),
    employed = "Employed",
    unemployed = "Unemployed",
    outside_labor_force = "Outside the labor force",
    all_working_age = "People 15 and older",
    labor_force = "Labor force",
    employed_people = "Employed people",
    as.character(value)
  )
}

domain_label <- function(value) {
  switch(
    as.character(value),
    national = "National",
    urban = "Urban",
    rural = "Rural",
    as.character(value)
  )
}

metric_value <- function(target_indicator, target_domain = "national", target_value = NULL) {
  row <- pick_row(target_indicator, target_domain, target_value)
  if (is.null(row)) {
    return(NA_real_)
  }
  as.numeric(row$display_estimate)
}

metric_count <- function(target_indicator, target_domain = "national", target_value = NULL) {
  row <- pick_row(target_indicator, target_domain, target_value)
  if (is.null(row)) {
    return(NA_real_)
  }
  as.numeric(row$display_estimate)
}

kpi_card <- function(label, value, note) {
  paste0(
    '<div class="labor-kpi">',
    "<strong>", html_escape(label), "</strong>",
    '<span class="labor-value">', html_escape(value), "</span>",
    "<p>", html_escape(note), "</p>",
    "</div>"
  )
}

bar_row <- function(label, value, max_value = 75) {
  width <- if (is.na(value)) 0 else max(0, min(100, as.numeric(value) / max_value * 100))
  paste0(
    '<div class="labor-bar-row">',
    "<span>", html_escape(label), "</span>",
    '<div class="labor-meter"><span style="width: ', sprintf("%.1f", width), '%;"></span></div>',
    '<strong style="text-align: right;">', fmt_pct(value, 1L), "</strong>",
    "</div>"
  )
}

condition_order <- c("employed", "unemployed", "outside_labor_force")

poverty_table_rows <- vapply(
  condition_order,
  function(value) {
    poverty <- metric_value("poverty_rate_by_labor_status", "national", value)
    extreme <- metric_value("extreme_poverty_rate_by_labor_status", "national", value)
    count <- metric_count("estimated_poor_by_labor_status", "national", value)
    paste0(
      "<tr>",
      "<td><strong>", html_escape(public_label(value)), "</strong></td>",
      "<td>", fmt_pct(poverty, 1L), "</td>",
      "<td>", fmt_pct(extreme, 1L), "</td>",
      "<td>", fmt_count(count), "</td>",
      "</tr>"
    )
  },
  character(1)
)

poverty_table <- paste0(
  '<div class="labor-table-wrap">',
  '<table class="table table-sm">',
  "<thead>",
  "<tr>",
  "<th>Labor condition</th>",
  "<th>Poverty rate</th>",
  "<th>Extreme poverty rate</th>",
  "<th>Estimated people in poverty</th>",
  "</tr>",
  "</thead>",
  "<tbody>",
  paste(poverty_table_rows, collapse = "\n"),
  "</tbody>",
  "</table>",
  "</div>"
)

structure_bar <- function(target_indicator, label) {
  value <- metric_value(target_indicator)
  bar_row(label, value, max_value = 75)
}

quality_bar <- function(target_indicator, label) {
  value <- metric_value(target_indicator)
  bar_row(label, value, max_value = 50)
}

poverty_bar <- function(value) {
  row <- pick_row("poverty_rate_by_labor_status", "national", value)
  if (is.null(row)) {
    return("")
  }
  bar_row(public_label(value), as.numeric(row$display_estimate), max_value = 40)
}

domain_summary_row <- function(domain_value) {
  paste0(
    "<tr>",
    "<td><strong>", domain_label(domain_value), "</strong></td>",
    "<td>", fmt_pct(metric_value("labor_force_participation_rate", domain_value), 1L), "</td>",
    "<td>", fmt_pct(metric_value("employment_rate_working_age", domain_value), 1L), "</td>",
    "<td>", fmt_pct(metric_value("unemployment_rate", domain_value), 1L), "</td>",
    "<td>", fmt_pct(metric_value("poverty_rate_by_labor_status", domain_value, "employed"), 1L), "</td>",
    "</tr>"
  )
}

domain_summary_table <- paste0(
  '<div class="labor-table-wrap">',
  '<table class="table table-sm">',
  "<thead>",
  "<tr>",
  "<th>Domain</th>",
  "<th>Labor participation</th>",
  "<th>Employment among people 15+</th>",
  "<th>Unemployment</th>",
  "<th>Poverty among employed people</th>",
  "</tr>",
  "</thead>",
  "<tbody>",
  paste(vapply(c("national", "urban", "rural"), domain_summary_row, character(1)), collapse = "\n"),
  "</tbody>",
  "</table>",
  "</div>"
)

period_label <- unique(as.character(labor$period))
if (length(period_label) != 1L) {
  period_label <- "2025"
}

participation <- metric_value("labor_force_participation_rate")
employment <- metric_value("employment_rate_working_age")
outside <- metric_value("outside_labor_force_share")
unemployment <- metric_value("unemployment_rate")
adequate <- metric_value("adequate_employment_rate")
subemployment <- metric_value("subemployment_rate")
employed_poverty <- metric_value("poverty_rate_by_labor_status", "national", "employed")
unemployed_poverty <- metric_value("poverty_rate_by_labor_status", "national", "unemployed")
outside_poverty <- metric_value("poverty_rate_by_labor_status", "national", "outside_labor_force")
employed_poor_count <- metric_count("estimated_poor_by_labor_status", "national", "employed")
outside_poor_count <- metric_count("estimated_poor_by_labor_status", "national", "outside_labor_force")
```

::: {.internal-hero .internal-hero-wide .labor-hero}
::: {.internal-hero-copy}
::: {.section-kicker}
Labor and income poverty
:::

# Labor market and poverty

::: {.story-lede}
Employment conditions and household vulnerability in annual ENEMDU 2025.
:::

Having work is central to household wellbeing, but employment alone does not guarantee that a household is above the income-poverty line. This page reads labor attachment, job quality and poverty together without treating one as a cause of the other.
:::

::: {.internal-hero-card .labor-scope-card}
## What this view covers

<dl class="labor-scope-list">
<div>
<dt>Period</dt>
<dd>Annual ENEMDU `r period_label`</dd>
</div>
<div>
<dt>Source</dt>
<dd>Curated aggregate site output</dd>
</div>
<div>
<dt>Coverage</dt>
<dd>National, urban and rural</dd>
</div>
<div>
<dt>Reading</dt>
<dd>Employment, job quality and poverty profiles</dd>
</div>
</dl>
:::
:::

::: {.content-block .editorial-flow}
::: {.story-panel}
::: {.section-kicker}
Public question
:::

## Why having work does not always mean being out of poverty

Poverty is measured through household per-capita income, not only through whether a person has a job. That distinction is crucial: a person can be employed and still live in a household whose total income is below the income-poverty line.

The national profile shows that poverty among employed people is **`r fmt_pct(employed_poverty, 1L)`**, while poverty among unemployed people is **`r fmt_pct(unemployed_poverty, 1L)`** and poverty among people outside the labor force is **`r fmt_pct(outside_poverty, 1L)`**. The estimated number of people in poverty is largest among employed people, at about **`r fmt_count(employed_poor_count)`**.
:::

::: {.labor-kpi-strip}
```{r}
#| echo: false
#| results: asis
cat(paste(
  kpi_card(
    "Labor-force participation",
    fmt_pct(participation, 1L),
    "Share of people age 15 and older in the labor force."
  ),
  kpi_card(
    "Employment among people 15+",
    fmt_pct(employment, 1L),
    "Share of people age 15 and older who are employed."
  ),
  kpi_card(
    "Outside the labor force",
    fmt_pct(outside, 1L),
    "Share of people age 15 and older outside the labor force."
  ),
  kpi_card(
    "Unemployment",
    fmt_pct(unemployment, 1L),
    "Share of the labor force that is unemployed."
  ),
  sep = "\n"
))
```
:::

::: {.story-grid}
::: {.story-panel}
::: {.section-kicker}
Labor-market structure
:::

## Who is attached to the labor market?

```{r}
#| echo: false
#| results: asis
cat(paste(
  '<div class="labor-bars">',
  structure_bar("labor_force_participation_rate", "Labor force participation"),
  structure_bar("employment_rate_working_age", "Employment among people 15+"),
  structure_bar("outside_labor_force_share", "Outside the labor force"),
  structure_bar("unemployment_rate", "Unemployment"),
  "</div>",
  sep = "\n"
))
```

The first reading separates labor-force attachment from unemployment. Many people are outside the labor force, while unemployment is measured only among people in the labor force.
:::

::: {.story-panel}
::: {.section-kicker}
Employment quality
:::

## Work status is not enough

```{r}
#| echo: false
#| results: asis
cat(paste(
  '<div class="labor-bars">',
  quality_bar("adequate_employment_rate", "Adequate/full employment"),
  quality_bar("subemployment_rate", "Subemployment"),
  "</div>",
  sep = "\n"
))
```

Among employed people, adequate/full employment reaches **`r fmt_pct(adequate, 1L)`** and subemployment reaches **`r fmt_pct(subemployment, 1L)`**. This is why the labor layer looks beyond the employed/unemployed split: job quality changes how employment should be read alongside household income poverty.
:::
:::

::: {.infographic-frame}
::: {.section-kicker}
Poverty by labor condition
:::

## Executive table

```{r}
#| echo: false
#| results: asis
cat(poverty_table)
```

The table compares broad labor conditions, not provinces or cities. The count column shows where the largest number of people in poverty is concentrated, while the poverty-rate columns show exposure within each labor condition.
:::

::: {.story-grid}
::: {.story-panel}
::: {.section-kicker}
Poverty profile
:::

## Poverty is not only a story about unemployment

```{r}
#| echo: false
#| results: asis
cat(paste(
  '<div class="labor-bars">',
  paste(vapply(condition_order, poverty_bar, character(1)), collapse = "\n"),
  "</div>",
  sep = "\n"
))
```

Unemployment is associated with a high poverty rate, but it accounts for a smaller estimated number of people in poverty than employment. The national estimate shows about **`r fmt_count(outside_poor_count)`** people in poverty outside the labor force, another reminder that household vulnerability extends beyond the labor-force headline.
:::

::: {.story-panel}
::: {.section-kicker}
Urban-rural reading
:::

## Labor and poverty differ by domain

```{r}
#| echo: false
#| results: asis
cat(domain_summary_table)
```

Urban and rural rows are included only for the supported annual domains in the curated output. They help compare labor attachment and poverty exposure without turning this page into a territorial ranking.
:::
:::

::: {.method-note}
## Reading notes

These estimates are analytical and descriptive. They are based on public ENEMDU annual 2025 microdata and are published here only through a curated aggregate output. They are not an official INEC validation, do not make causal claims, and should be read together with [Measurement and statistical quality](04_methodology-quality.qmd).
:::

::: {.proof-strip}
::: {.proof-item}
**Income poverty**

Read the household-income threshold behind the poverty measure.
:::

::: {.proof-item}
**Territory**

Compare incidence and concentration across provinces.
:::

::: {.proof-item}
**Labor**

Read employment, job quality and household vulnerability together.
:::

::: {.proof-item}
**Measurement**

Keep domains, uncertainty and publication limits visible.
:::
:::

::: {.page-nav}
<a href="02_income-poverty.qmd">Income poverty</a>
<a href="05_territorial-province-income-poverty.qmd">Territorial poverty</a>
<a href="04_methodology-quality.qmd">Measurement and statistical quality</a>
:::
:::
